---
title: "Truly Sparse Multivariate Regression"
author: "Sean Corum"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The purpose of the Truly Sparse Multivariate Regression (*tsmvr*) package is to solve linear multivariate regression problems. A multivariate regression problem is a regression problem with more than one response. The form of such a problem is

\begin{equation}
 \mathbf{Y} = \mathbf{X} \mathbf{B} + \mathbf{E} .
\end{equation}

Here, $\mathbf{X}$ is the design matrix encoding $n$ observations of $p$ features, $\mathbf{Y}$ is the response matrix encoding $n$ observations of $q$ responses, $\mathbf{B}$ is the regressor matrix, and $\mathbf{E}$ is the error term. Given $\mathbf{X}$ and $\mathbf{Y}$, $\mathbf{B}$ is solved for while also simultaneously estimating the error covariance. During this process, the sparsity of $\mathbf{B}$ and the covariance is strictly enforced. This yields higher quality solutions, especially for problems with true sparse solutions.

The algrorithm in *tsmvr* is adapted from "High Dimensional Multivariate Regression and Precision Matrix Estimation via Nonconvex Optimization" by Chen and Gu (arXiv:1606.00832, 2016).

## Examples

### An Introductory Example
```{r intro_data}
library(tsmvrextras)
n = 50000
p = 5000
q = 30
sparsity = 0.1
s1 = p * q * 0.2
s2 = 3*q
```

```{r intro_solve}
data = tsmvrextras::make_data(
  n = n, p=p, q=q, b1 = sqrt(sparsity), b2 = sqrt(sparsity), 
  rep = 1
)[[1]]
solution = tsmvr::tsmvr_solve(
  X = data$X, Y = data$Y, s1 = s1, s2 = s2, B_type = 'gd', 
  Omega_type = 'gd', eta1 = 0.2, eta2 = 1e-5,
  max_iter = 2000, skip = 1
)
```

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))

# References


